Comparing clustering parameters with rOpenScPCA

### 2025-01-14 analyses/hello-clusters/02_compare-clustering-parameters.Rmd
swap in new rprojroot line so this runs in CI, and re-render notebooks
Apply suggestions from code review
Co-authored-by: Joshua Shapiro <josh.shapiro@ccdatalab.org>

This function takes a PCA matrix with row names representing unique cell ids (e.g., barcodes) as its primary argument, with additional arguments for cluster parameters. 
This function wraps the `rOpenScPCA::calculate_clusters()` function but allows you to provide a vector of parameter values to perform clustering across, as listed below.
Clusters will be calculated for all combinations of parameters values (where applicable); default values that the function will use for any unspecified parameter values are shown in parentheses

@@

```{r sweep clusters}
# Define nn parameter values of interest
nn_values <- seq(10, 30, 10)

@@

      # calculate neighbhorhood purity
      purity_df <- rOpenScPCA::calculate_purity(pca_matrix, cluster_df)

      # Combine into a single data frame
      dplyr::left_join(silhouette_df, purity_df)
    }
  )

@@

```{r calculate stability}
stability_list <- cluster_results_list |>
  purrr::map(
    \(cluster_df) {
      nn <- unique(cluster_df$nn)

@@

```{r sweep two parameters}
# Define vectors of parameters to vary
nn_values <- seq(10, 30, 10)
res_values <- seq(5, 15, 5) / 10

@@

@@

```{r visualize purity two parameters}
ggplot(purity_df) +
  aes(x = as.factor(nn), y = purity, fill = as.factor(nn)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Pastel2") +
  # facet by resolution
  facet_wrap(vars(resolution_label)) +
  labs(
    x = "Number of nearest neighbors",
    y = "Neighborhood purity"
  ) +
  theme(legend.position = "none")
```

>>>>

This function takes a PCA matrix with row names representing unique cell ids (e.g., barcodes) as its primary argument, with additional arguments for cluster parameters. 
This function wraps the `rOpenScPCA::calculate_clusters()` function but allows you to provide a vector of parameter values to perform clustering across, as listed below.
Clusters will be calculated for all combinations of parameters values (where applicable); default values that the function will use for any unspecified parameter values are shown in parentheses.

@@

```{r sweep clusters}
# Define nn parameter values of interest
nn_values <- c(10, 20, 30)

@@

      # calculate neighbhorhood purity
      purity_df <- rOpenScPCA::calculate_purity(pca_matrix, silhouette_df)
    }
  )

@@

```{r calculate stability}
stability_list <- cluster_results_list |>
  purrr::map(
    \(cluster_df) {
      nn <- cluster_df$nn[1] # all rows have the same `nn` parameter, so we'll take the first

@@

```{r sweep two parameters}
# Define vectors of parameters to vary
nn_values <- c(10, 20, 30)
res_values <-c(0.5, 1.0, 1.5)

@@

# Print the plots with patchwork::wrap_plots()
patchwork::wrap_plots(umap_plots, ncol = 3)
```

@@

```{r visualize purity two parameters}
ggplot(purity_df) +
  aes(x = as.factor(nn), y = purity, fill = as.factor(nn)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Pastel2") +
  # facet by resolution
  facet_wrap(vars(resolution_label), labeller = label_both) +
  labs(
    x = "Number of nearest neighbors",
    y = "Neighborhood purity"
  ) +
  theme(legend.position = "none")
```
